Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

-------------------------------------------------------------------------
Here I use basic R command to learn to use this new language!!!!

Creating a vector:

```{r}
x <-c(1,3,2,5)  #c-->concatenate
y <-c(1,4,3)
```
Check the length of the vectors
```{r}
length(x)
length(y)
```
List of objects / delete objects
```{r}
ls()  #list of all objects
rm(x,y) #delete objects, we can also use rm(list=ls()) to delete all objects
```
Documentation about a function

?FunctionName
```{r}
?matrix
```
Working with matrixes
```{r}
x = matrix(data=c(1,2,3,4),nrow=2,ncol=2) 
#We can also write: x=matrix(data=c(1,2,3,4),2,2)
y = matrix(data=c(1,2,3,4),nrow=2,ncol=2,byrow=TRUE)  #by default R create a matrix by column

#Operations with matrixes
sqrt(x)
x^2
```
Working with random variables (basic concepts)
```{r}
x = rnorm(50)  #creates a vector of 50 values from a standard normal distribution
y = x + rnorm(50,mean=50,sd=.1) #we add extract random variables from a distribution with a specific shape (specific mean and std deviation)


```








