Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

-------------------------------------------------------------------------
Here I use basic R command to learn to use this new language!!!!

Creating a vector:

```{r}
x <-c(1,3,2,5)  #c-->concatenate
y <-c(1,4,3)
```
Check the length of the vectors
```{r}
length(x)
length(y)
```
List of objects / delete objects
```{r}
ls()  #list of all objects
rm(x,y) #delete objects, we can also use rm(list=ls()) to delete all objects
```
Documentation about a function

?FunctionName
```{r}
?matrix
```
Working with matrixes
```{r}
x = matrix(data=c(1,2,3,4),nrow=2,ncol=2) 
#We can also write: x=matrix(data=c(1,2,3,4),2,2)
y = matrix(data=c(1,2,3,4),nrow=2,ncol=2,byrow=TRUE)  #by default R create a matrix by column

#Operations with matrixes
sqrt(x)
x^2
```
Working with random variables (basic concepts)
```{r}
x = rnorm(50)  #creates a vector of 50 values from a standard normal distribution
y = x + rnorm(50,mean=50,sd=.1) #we add extract random variables from a distribution with a specific shape (specific mean and std deviation)
#Use set.seed(3)-->thave same random variables (with same length vector)
var(y)
mean(y)
sd(y) # =sqrt(var(y))
```
Working with graphics
```{r}
x = rnorm(100)
y = rnorm(100)
jpeg("FiguraTest.jpeg")  #We create a jpeg of the plot
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="PLot of X vs Y")
dev.off()
```
More sophisticate plot
```{r}
x = seq(1,10) #a sequence of number from 1 to 10
#x=1:10 is equivalent

y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
"contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2
contour (x,y,fa,nlevels =15)"
```
```{r}
x = seq(1,1000)
y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
image(x,y,f) #produce a heatmap
```
Plotting a three-dimensional plot
```{r}
x = seq(1,100)
y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
persp(x,y,f,theta=30,phi=30) #theta and phi control the angles at which the plot is viewed
```
Indexing Data
```{r}
A = matrix(1:16,4,4)
A[2,3]
```


```{r}
A[c(1,3),c(2,4)]  #slicing a matrix
A[1:3,2:4]  #Another way...
A[1:2,]
A[,1:2]
```
```{r}
A[-c(1,3),]  #don't consider rows 1 and 3
dim(A) #shape of the matrix (rows, columns)
```
Loading Data
```{r}
# library(ISLR) -->installing a library
Auto #this dataset is already present when we installed ISLR library
names(Auto) #to see columns name
```
Additional Graphical and Numerical Summaries
```{r}
plot(Auto$cylinders,Auto$mpg)
#attach(Auto) so we can do:
#plot(cylinders,mpg)
#remember!!!: detach(Auto)
```
```{r}
Auto$cylinders=as.factor(Auto$cylinders)  #transform Auto from numeric to categorical (qualitative var)
```
```{r}
plot(Auto$cylinders,Auto$mpg)  #boxplot is printed if  i have categorical variable on x-axis
```
```{r}
hist(Auto$mpg[Auto$cylinders=="4"])  #Added a filter to histogram
```
```{r}
pairs( ~mpg + cylinders + horsepower,Auto)  #produce a scatterplot among the combination of all features
```
```{r}
summary(Auto) #a summary of Auto dataset
summary(Auto$weight) #summary of a single attribute
```

```{r}
savehistory()
#use loadhistory to load history next time
```
###########EXERCISES#############
```{r}
College
```
```{r}
names(College)
```
```{r}
College
```
```{r}
summary(College)
```
```{r}
pairs(College[,1:10])  #scatterplot of first 10 features (columns)
```
```{r}
plot(College$Outstate,College$Private)
```
We want to create a qualitative variable [Yes/No] with college$Top10perc > or < 50
```{r}
Elite = rep("No",nrow(College))
Elite[College$Top10perc>50]="Yes"
Elite=as.factor(Elite)
College=data.frame(College,Elite)
```


```{r}
College
```
```{r}
summary(College$Elite)
```
```{r}
par(mfrow=c(2,2)) #create a 2x2 plot
hist(College$Accept)
hist(College$Enroll)
hist(College$Top10perc)
hist(College$Top25perc)
```
A function to set image size:
```{r}
Auto
```
Finding min/max of quantitative variables

```{r}
Auto=na.omit(Auto) #removes rows with missing values
names(Auto)
```


```{r}
range(Auto$displacement)
range(Auto$mpg)
range(Auto$horsepower)
range(Auto$weight)
range(Auto$acceleration)
range(Auto$year)
range(Auto$origin)
```
Mean and standard deviation of quantitative variables
```{r}
print("####Mean###")
mean(Auto$displacement)
mean(Auto$mpg)
mean(Auto$horsepower)
mean(Auto$weight)
mean(Auto$acceleration)
mean(Auto$year)
mean(Auto$origin)
print("####Variance:####")
sqrt(var(Auto$displacement))
sqrt(var(Auto$mpg))
sqrt(var(Auto$horsepower))
sqrt(var(Auto$weight))
sqrt(var(Auto$acceleration))
sqrt(var(Auto$year))
sqrt(var(Auto$origin))

```
Slicing Auto dataset
```{r}
Auto1 = Auto[-c(rep(10:85)),]  #We remove from the dataset rows from 10 to 85
```
We recalculate Min/Max, Mean, Std
```{r}
print("####Min/Max####")
range(Auto1$displacement)
range(Auto1$mpg)
range(Auto1$horsepower)
range(Auto1$weight)
range(Auto1$acceleration)
range(Auto1$year)
range(Auto1$origin)
print("####Mean###")
mean(Auto1$displacement)
mean(Auto1$mpg)
mean(Auto1$horsepower)
mean(Auto1$weight)
mean(Auto1$acceleration)
mean(Auto1$year)
mean(Auto1$origin)
print("####Variance:####")
sqrt(var(Auto1$displacement))
sqrt(var(Auto1$mpg))
sqrt(var(Auto1$horsepower))
sqrt(var(Auto1$weight))
sqrt(var(Auto1$acceleration))
sqrt(var(Auto1$year))
sqrt(var(Auto1$origin))
```
Plotting scatterplots for combination of different predictors(features)
```{r}
pairs(~ mpg + displacement + horsepower+ cylinders+weight +acceleration 
 , Auto)
```







