Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

-------------------------------------------------------------------------
Here I use basic R command to learn to use this new language!!!!

Creating a vector:

```{r}
x <-c(1,3,2,5)  #c-->concatenate
y <-c(1,4,3)
```
Check the length of the vectors
```{r}
length(x)
length(y)
```
List of objects / delete objects
```{r}
ls()  #list of all objects
rm(x,y) #delete objects, we can also use rm(list=ls()) to delete all objects
```
Documentation about a function

?FunctionName
```{r}
?matrix
```
Working with matrixes
```{r}
x = matrix(data=c(1,2,3,4),nrow=2,ncol=2) 
#We can also write: x=matrix(data=c(1,2,3,4),2,2)
y = matrix(data=c(1,2,3,4),nrow=2,ncol=2,byrow=TRUE)  #by default R create a matrix by column

#Operations with matrixes
sqrt(x)
x^2
```
Working with random variables (basic concepts)
```{r}
x = rnorm(50)  #creates a vector of 50 values from a standard normal distribution
y = x + rnorm(50,mean=50,sd=.1) #we add extract random variables from a distribution with a specific shape (specific mean and std deviation)
#Use set.seed(3)-->thave same random variables (with same length vector)
var(y)
mean(y)
sd(y) # =sqrt(var(y))
```
Working with graphics
```{r}
x = rnorm(100)
y = rnorm(100)
jpeg("FiguraTest.jpeg")  #We create a jpeg of the plot
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="PLot of X vs Y")
dev.off()
```
More sophisticate plot
```{r}
x = seq(1,10) #a sequence of number from 1 to 10
#x=1:10 is equivalent

y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
"contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2
contour (x,y,fa,nlevels =15)"
```
```{r}
x = seq(1,1000)
y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
image(x,y,f) #produce a heatmap
```
Plotting a three-dimensional plot
```{r}
x = seq(1,100)
y = x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
persp(x,y,f,theta=30,phi=30) #theta and phi control the angles at which the plot is viewed
```






